DeepSphere: Efficient spherical Convolutional Neural Network with HEALPix sampling for cosmological applications

### 2019-03-26 scnn.tex
cite PyGSP
architecture figure: spatial summarization

We give a practical demonstration of DeepSphere on cosmological model discrimination using maps of projected mass distribution on the sky \citep{chang2017curvedsky}.
These kind of maps can be created using the gravitational lensing technique \citep[see][for review]{BartelmannSchneider2001weak}.
Our maps are similar to the ones used by \citep{schmelze2017cosmologicalmodel}.
In a simplified scenario, we classify partial sky convergence maps into two cosmological models.
These models were designed to have very similar angular power spectrum.
We compare the performance of DeepSphere to three baselines: a 2D CNN, and an SVM classifier that takes pixel histograms or power spectral densities (PSDs) of these maps as input.
The comparison is made as a function of the additive noise level and the area of the sphere used in the analysis.
Results show that our model is always better at discriminating the maps, especially in the presence of noise.
DeepSphere is implemented with \pkg{TensorFlow} \citep{abadi2016tensorflow} and is intended to be easy to use out-of-the-box for cosmological applications.
Code and examples are available online.{\interfootnotelinepenalty=10000 \footnote{\url{https://github.com/SwissDataScienceCenter/DeepSphere/}}}

@@

\begin{figure}[t!]
	\centering
	\includegraphics[width=\linewidth]{subspace_harmonics_eigenvectors_v2}
	\caption{Correspondence between the subspaces spanned by the graph Fourier modes and the spherical harmonics.
		%The correspondence is computed in two steps.
		First, we compute the power spectral density (PSD) of each graph eigenvector.
		Second, as there is $2\ell+1$ spherical harmonics for each degree $\ell$, we sum the contributions of the corresponding $2\ell+1$ graph eigenvectors.
		% No real convergence with increasing Nside.
		% Row $j$ is the sum of the (normalized) power spectral densities (PSDs), computed by the SHT, of the $2j+1$ graph Fourier modes $\b u_i$ identified with degree $j$.
		% (\figref{graph_eigenvalues} shows how Fourier modes can be identified to spherical degrees.)
		% The PSD is computed for degrees $\ell=0$ to $\ell = 3 N_{side} -1$.
		% Pixel $k$ in row $j$ measures the energy in frequency band $k$ of the graph Fourier modes identified to be part of band $j$.
		The matrix shows how the subspaces agree: the subspaces spanned by the Fourier modes correspond to those spanned by the spherical harmonics in the low frequencies, and the Fourier modes leak towards adjacent frequency bands at higher frequencies.
		While there is a systematic error, the Fourier modes agree at higher frequencies as $N_{side}$ increases.}
		\label{fig:subspace_harmonics_eigenvectors}
\end{figure}

>>>>

We give a practical demonstration of DeepSphere on cosmological model discrimination using maps of projected mass distribution on the sky \citep{chang2017curvedsky}.
These kind of maps can be created using the gravitational lensing technique \citep[see][for review]{BartelmannSchneider2001weak}.
Our maps are similar to the ones used by \citep{schmelze2017cosmologicalmodel}.
In a simplified scenario, we classify partial sky convergence maps into two cosmological models.
These models were designed to have very similar angular power spectrum.
We compare the performance of DeepSphere to three baselines: a 2D CNN, and an SVM classifier that takes pixel histograms or power spectral densities (PSDs) of these maps as input.
The comparison is made as a function of the additive noise level and the area of the sphere used in the analysis.
Results show that our model is always better at discriminating the maps, especially in the presence of noise.
DeepSphere is implemented with \pkg{TensorFlow} \citep{abadi2016tensorflow} and is intended to be easy to use out-of-the-box for cosmological applications.
Many plots and co PyGSP \citep{pygsp} for computations and plots.
The Python Graph Signal Processing package (PyGSP) \citep{pygsp} is used to build graphs, compute the Laplacian and Fourier basis, and perform graph convolutions.
Code and examples are available online.{\interfootnotelinepenalty=10000 \footnote{\url{https://github.com/SwissDataScienceCenter/DeepSphere/}}}

@@

\begin{figure}[t!]
	\centering
	\includegraphics[width=\linewidth]{subspace_harmonics_eigenvectors_v2}
	\caption{Correspondence between the subspaces spanned by the graph Fourier modes and the spherical harmonics.
		%The correspondence is computed in two steps.
		First, we compute the power spectral density (PSD) of each graph eigenvector with the SHT.
		Second, as there is $2\ell+1$ spherical harmonics for each degree $\ell$, we sum the contributions of the corresponding $2\ell+1$ graph eigenvectors.
		% No real convergence with increasing Nside.
		% Row $j$ is the sum of the (normalized) power spectral densities (PSDs), computed by the SHT, of the $2j+1$ graph Fourier modes $\b u_i$ identified with degree $j$.
		% (\figref{graph_eigenvalues} shows how Fourier modes can be identified to spherical degrees.)
		% The PSD is computed for degrees $\ell=0$ to $\ell = 3 N_{side} -1$.
		% Pixel $k$ in row $j$ measures the energy in frequency band $k$ of the graph Fourier modes identified to be part of band $j$.
		The matrix shows how the subspaces align: the Fourier modes span the same subspaces as the spherical harmonics in the low frequencies, and the Fourier modes leak towards adjacent frequency bands at higher frequencies.
		While there is a systematic error, the Fourier modes align at higher frequencies as $N_{side}$ increases.}
		\label{fig:subspace_harmonics_eigenvectors}
\end{figure}
